---
title: "20201110-p8119_hw3_jsg2145"
author: "Jared Garfinkel"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(SKAT)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
tibble(Chapter = c(7, 10, 11),
       Problems = c("9", "2, 3", "3"))
```

# Chapter 7

## Problem 9

```{r}
data = read_delim("./data/Takei_SNPs.txt", delim = " ")
data
```

To test $H_0$: $p_{cases}$ = $p_{controls}$ without assuming HWE,

```{r}
summary_table = data %>% 
  group_by(SNPID) %>% 
  summarize(r_M = 2*MM+Mm,
            r_m = 2*mm+Mm,
            s_M = 2*MM_1 + Mm_1,
            s_m = 2*mm_1 + Mm_1,
            n_M = 2*(MM + MM_1) + (Mm + Mm_1),
            n_m = 2*(mm + mm_1) + (Mm + Mm_1),
            r = MM + Mm + mm,
            s = MM_1 + Mm_1 + mm_1,
            n = MM + Mm + mm + MM_1 + Mm_1 + mm_1,
            xcases = (2*MM + Mm)/r,
            xcontrols = (2*MM_1 + Mm_1)/s,
            pcases = xcases/2,
            pcontrols = xcontrols/2,
            x = (xcases + xcontrols)/2,
            p = (pcases + pcontrols)/2,
            zt = 2*sqrt(r*s)*(pcases - pcontrols)/sqrt(4*(MM+MM_1)+(Mm+Mm_1)-4*n*p^2))
# zt = 2*sqrt(pcases -pcontrol)
```


```{r}
OR_tab = data %>% 
  summarize(MM = sum(MM),
            Mm = sum(Mm),
            mm = sum(mm),
            MM_1 = sum(MM_1),
            Mm_1 = sum(Mm_1),
            mm_1 = sum(mm_1))
tibble(Status =
         c("Cases", "Controls"),
       `Exposed (r1)` = 
         c(Disease = pull(OR_tab, Mm),
                Control = pull(OR_tab, Mm_1)),
       `Unexposed (r0)` = 
         c(Disease = pull(OR_tab, MM), 
                Control = pull(OR_tab, MM_1)))
  
  # unnest(cols = c(Exposed, Unexposed))
```

$OR_{r_1}$ = a * d / b * c = `r round(6300 * 16053 /6983 / 10581, 3)`

```{r}
tibble(Status =
         c("Cases", "Controls"),
       `Exposed (r2)` = 
         c(Disease = pull(OR_tab, mm),
                Control = pull(OR_tab, mm_1)),
       `Unexposed (r0)` = 
         c(Disease = pull(OR_tab, MM), 
                Control = pull(OR_tab, MM_1)))
```

$OR_{r_2}$ = `r round(1876 * 16053 / 1451 / 10581, 3)`

# Chapter 10

## Problem 2

Compute the Bonferroni correction for the SNPs.

There are 35 SNPs.

So, we divide the significance level by 35.

$\alpha \le M*\alpha'$ = `r round(0.05/35, 5)`

$Z_T \le \phi(0.00143)$

```{r}
# qnorm(1-0.00143)
summary_table %>% 
  mutate(zt2 = zt^2,
         chisq = qchisq(1-0.00143, 1)) %>% 
  filter(abs(zt2)>chisq)
```

It appears that all the SNPs are significant even using the Bonferroni adjustment.

## Problem 3

```{r}
summary_table %>% 
  mutate(p_zt = dnorm(zt)) %>% 
  arrange(p_zt) %>% 
  mutate(rank_p_zt = row_number(),
         FDR = rank_p_zt/35*0.05) %>% 
  filter(p_zt < FDR)
```

All the p-values are significant using a modified Simes-procedure.

For all (i), $p_i < \frac{(i)}{M}\alpha$

# Chapter 11

## Problem 3

$1 - \alpha$ = P(for i SNPs, $H_0$ is not rejected when the null is true)

$$1 - \alpha = \sum_{i = 0}^{M}\left[{M \choose i} (\alpha_1(1-\alpha_2))^i(1-\alpha_1)^{m-i}\right]$$

= $$[\alpha_1(1-\alpha_2) + (1-\alpha_1)]^M$$

Since $(1-\alpha)^\frac{1}{M}$ ~ $1 - \frac{\alpha}{M}$,

$$\frac{\alpha}{M} = \alpha_1*\alpha_2*...*\alpha_K$$

Thus, $$\frac{\alpha}{M} = \prod_{k = 1}^{M}\alpha_i$$

# SKAT

```{r}
data(SKAT.example)
names(SKAT.example)
attach(SKAT.example)
```

```{r}
skimr::skim(SKAT.example)
```

There are 67 variables of Z - Z variables are mostly 0, but range from 0 to 2.

In the variables X.1 and Y.b there are an equal number of observations taking values of 0 and 1.

X.2 and y.c appear to be continuous variables centered around 0 with a a variance near 1.

```{r}
df = SKAT.example %>%
  data.frame() %>% 
  tibble() %>% 
  janitor::clean_names()
```


```{r}
df %>% 
  ggplot(aes(x = x_2, fill = as_factor(x_1))) +
  geom_histogram(alpha = 0.3)
```

```{r}
df %>% 
  ggplot(aes(x = y_c, fill = as_factor(y_b))) +
  geom_histogram(alpha = 0.4)
```


It appears that the grouping variables X.1 and y.b do not overlap.

```{r}
obj<-SKAT_Null_Model(y.c ~ X, out_type="C", data = SKAT.example)
```

```{r}
#SKAT test
SKAT(Z, obj, r.corr=0)$p.value
```


```{r}
#Burden test
SKAT(Z, obj,r.corr=1)$p.value
```


```{r}
# SKAT-O: Optimal over r.corr in [0,1]; r.corr is the 
# correlation among the beta’s of individual variants in 
# the gene.
SKAT(Z, obj, method="optimal")$p.value
```

